```{r}
setwd('/Users/v.baranov/Desktop/Диплом/code/R')
library(biomod2)
library(ggplot2)
library(gridExtra)
library(raster)
library(rasterVis)
library(rgeos)
library(rgdal)
library(sp)
```

# Read data with thermokarst (which was previously converted to Jupyter)
```{r}
termokarst_occ <- read.csv('data/termo_points_2.csv')
summary(termokarst_occ)
```

# Create a stack with the necessary variables
```{r}
WС_sub_1 <- 
  raster::stack(
    c(
      bio_1 = 'data/worldclim/bio_2.5/rasters/clip_wc2.1_2.5m_bio_1.tif',
      bio_5 = 'data/worldclim/bio_2.5/rasters/clip_wc2.1_2.5m_bio_5.tif',
      bio_7 = 'data/worldclim/bio_2.5/rasters/clip_wc2.1_2.5m_bio_7.tif',
      bio_10 = 'data/worldclim/bio_2.5/rasters/clip_wc2.1_2.5m_bio_10.tif',
      bio_12 = 'data/worldclim/bio_2.5/rasters/clip_wc2.1_2.5m_bio_12.tif',
      bio_18 = 'data/worldclim/bio_2.5/rasters/clip_wc2.1_2.5m_bio_18.tif',
      elev = 'data/worldclim/dem_2.5/elev_2.5_clipped.tif'
    )
  )

```

## Formatting data for the model
```{r}
termokarst_data <- 
  BIOMOD_FormatingData(
    expl.var = WС_sub_1,
    resp.var = termokarst_occ['TKThLP'],
    resp.xy = termokarst_occ[, c('lon', 'lat')],
    resp.name = "TO",
    # PA.nb.rep = 1,
    # PA.nb.absences = 100,
    # PA.strategy = 'random'
  )

plot(termokarst_data)
```
```{r}
termokarst_data
```

# Individual BIOMOD2 models settings
```{r}
termo_opt <- 
  BIOMOD_ModelingOptions(
    GLM = list(type = 'quadratic', interaction.level = 1),
    GBM = list(n.trees = 1000),
    GAM = list(algo = 'GAM_mgcv')
  )
```

# Modelling
```{r}
termo_models_1 <-  
  BIOMOD_Modeling(
    data = termokarst_data,
    models = c("GLM", "GBM", "RF", "GAM"),
    models.options = termo_opt,
    NbRunEval = 2, #два прогона
    DataSplit = 80,
    VarImport = 3,
    modeling.id = "first_try"
  )
```

## Load model's evaluations

```{r}
termo_models_1_scores <- get_evaluations(termo_models_1)
```

## 5 dimension array containing the scores of the models

```{r} 
dim(termo_models_1_scores)
dimnames(termo_models_1_scores)
```

## Model's graphs
```{r}
models_scores_graph(
  termo_models_1, 
  by = "models", 
  metrics = c("ROC","TSS"), 
  xlim = c(0.5,1), 
  ylim = c(0.5,1)
)
#TSS - true skill statistics
#An ROC curve (receiver operating characteristic curve) is a graph showing the performance of a classification model at all classification thresholds

models_scores_graph(
  termo_models_1, 
  by = "cv_run" , 
  metrics = c("ROC","TSS"), 
  xlim = c(0.5,1), 
  ylim = c(0.5,1)
)

models_scores_graph(
  termo_models_1, 
  by = "data_set", 
  metrics = c("ROC","TSS"), 
  xlim = c(0.5,1), 
  ylim = c(0.5,1)
)
```

## Variables importance
```{r}
(termo_models_1_var_import <- get_variables_importance(termo_models_1))
```

# Average "importance" for all runs
```{r}
apply(termo_models_1_var_import, c(1,2), mean)
```

## Response curves for each model
```{r}
termo_glm <- BIOMOD_LoadModels(termo_models_1, models='GLM')
termo_gbm <- BIOMOD_LoadModels(termo_models_1, models='GBM')
termo_rf <- BIOMOD_LoadModels(termo_models_1, models='RF')
termo_gam <- BIOMOD_LoadModels(termo_models_1, models='GAM')

glm_eval_strip <- 
  biomod2::response.plot2(
    models  = termo_glm,
    Data = get_formal_data(termo_models_1,'expl.var'), 
    show.variables= get_formal_data(termo_models_1,'expl.var.names'),
    do.bivariate = FALSE,
    fixed.var.metric = 'median',
    legend = FALSE,
    data_species = get_formal_data(termo_models_1,'resp.var')
  )

gbm_eval_strip <- 
  biomod2::response.plot2(
    models  = termo_gbm,
    Data = get_formal_data(termo_models_1,'expl.var'), 
    show.variables= get_formal_data(termo_models_1,'expl.var.names'),
    do.bivariate = FALSE,
    fixed.var.metric = 'median',
    legend = FALSE,
    data_species = get_formal_data(termo_models_1,'resp.var')
  )

rf_eval_strip <- 
  biomod2::response.plot2(
    models  = termo_rf,
    Data = get_formal_data(termo_models_1,'expl.var'), 
    show.variables= get_formal_data(termo_models_1,'expl.var.names'),
    do.bivariate = FALSE,
    fixed.var.metric = 'median',
    legend = FALSE,
    data_species = get_formal_data(termo_models_1,'resp.var')
  )

gam_eval_strip <- 
  biomod2::response.plot2(
  models  = termo_gam,
  Data = get_formal_data(termo_models_1,'expl.var'), 
  show.variables= get_formal_data(termo_models_1,'expl.var.names'),
  do.bivariate = FALSE,
  fixed.var.metric = 'median',
  # legend = FALSE,
  data_species = get_formal_data(termo_models_1,'resp.var')
)
```

# Ensemble models
```{r}
termo_ensemble_models_1 <- 
  BIOMOD_EnsembleModeling(
    modeling.output = termo_models_1,
    em.by = 'all',
    eval.metric = 'TSS',
    eval.metric.quality.threshold = 0.6, 
    models.eval.meth = c('TSS','ROC'),
    prob.mean = FALSE,
    prob.cv = TRUE, 
    committee.averaging = TRUE,
    prob.mean.weight = TRUE,
    VarImport = 0
  )

```

## Ensemble models score
```{r}
(termo_ensemble_models_1_scores <- get_evaluations(termo_ensemble_models_1))
#cv - coefficient variation
#PA - pseudo absences
#ca - comute average
#wm - Weighted mean
```

# Do models projections

## current projections
```{r}
termo_models_1_proj_current <- 
  BIOMOD_Projection(
    modeling.output = termo_models_1,
    new.env = WС_sub_1,
    proj.name = "current",
    binary.meth = "TSS",
    output.format = ".img",
    do.stack = FALSE
  )

termo_ensemble_models_1_proj_current <- 
  BIOMOD_EnsembleForecasting(
    EM.output = termo_ensemble_models_1,
    projection.output = termo_models_1_proj_current,
    binary.meth = "TSS",
    output.format = ".img",
    do.stack = FALSE
  )
```

## future projections
```{r}
WС_2040_sub_1 <- 
   raster::stack(
    c(
      bio_1 = 'data/worldclim/access_ssp126_2021_2040/clipped/rasters/clip_World_1.tif',
      bio_5 = 'data/worldclim/access_ssp126_2021_2040/clipped/rasters/clip_World_5.tif',
      bio_7 = 'data/worldclim/access_ssp126_2021_2040/clipped/rasters/clip_World_7.tif',
      bio_10 = 'data/worldclim/access_ssp126_2021_2040/clipped/rasters/clip_World_10.tif',
      bio_12 = 'data/worldclim/access_ssp126_2021_2040/clipped/rasters/clip_World_12.tif',
      bio_18 = 'data/worldclim/access_ssp126_2021_2040/clipped/rasters/clip_World_18.tif',
      elev = 'data/worldclim/dem_2.5/elev_2.5_clipped.tif'
    )
  )

termo_models_1_proj_2040 <- 
  BIOMOD_Projection(
    modeling.output = termo_models_1,
    new.env = WС_2040_sub_1,
    proj.name = "termo_2040",
    binary.meth = "TSS",
    output.format = ".img",
    do.stack = FALSE
  )

termo_ensemble_models_1_proj_2040 <- 
  BIOMOD_EnsembleForecasting(
    EM.output = termo_ensemble_models_1,
    projection.output = termo_models_1_proj_2040,
    binary.meth = "TSS",
    output.format = ".img",
    do.stack = FALSE
  )
```

```{r}
plot(termo_ensemble_models_1_proj_2040
     , str.grep = "EMca|EMwmean"
     )
```

## Calculating SRC (species range change)
## load models
```{r}
term_bin_proj_current <- 
  stack( 
    c(
      ca = "TO/proj_current/individual_projections/TO_EMcaByTSS_mergedAlgo_mergedRun_mergedData_TSSbin.img",
      wm = "TO/proj_current/individual_projections/TO_EMwmeanByTSS_mergedAlgo_mergedRun_mergedData_TSSbin.img"
    )
  )

termo_bin_proj_2040 <- 
  stack( 
    c(
      ca = "TO/proj_termo_2040/individual_projections/TO_EMcaByTSS_mergedAlgo_mergedRun_mergedData_TSSbin.img",
      wm = "TO/proj_termo_2040/individual_projections/TO_EMwmeanByTSS_mergedAlgo_mergedRun_mergedData_TSSbin.img"
    )
  )

SRC_current_2040 <- 
  BIOMOD_RangeSize(
    term_bin_proj_current,
    termo_bin_proj_2040
  )

SRC_current_2040$Compt.By.Models
```
```{r}
SRC_current_2040$Diff.By.Pixel
```
## Colorkey for SRC
```{r}
names(SRC_current_2040$Diff.By.Pixel) <- c("ca cur-2040", "wm cur-2040")

my.at <- seq(-2.5, 1.5, 1)
myColorkey <- 
  list(
    at = my.at, ## where the colors change
    labels = 
      list(
        labels = c("lost", "pres", "abs","gain"), ## labels
        at = my.at[-1] - 0.5 ## where to print labels
      )
  )

```
## Coloring SRC map

```{r}
# tiff(filename = 'output.tiff')
rasterVis::levelplot( 
  SRC_current_2040$Diff.By.Pixel, 
  main = "Thermokarst lakes range change",
  colorkey = myColorkey,
  col.regions=c('#f03b20', '#99d8c9', '#f0f0f0', '#2ca25f'),
  # layout = c(2,2)
)
# dev.off()
```

```{r}
SRC_current_2040
```
## Saving results as TIFF 
```{r}
future_2040 <- brick(SRC_current_2040$Diff.By.Pixel)
for (i in 1:nlayers(future_2040))
{
proj_2040 <-(future_2040[[i]])
writeRaster(proj_2040, glue::glue("/Users/v.baranov/Desktop/Диплом/code/R/results/25_05_22/proj_2040{i}.tif"), sep="")
}
```